proof-of-concept_mods

### 2023-11-06 proof-of-concept_mods.Rmd
comparing abs centile diffs w/in pheno, troubleshooting missing cf-cf.lm t-test results

Test for differences in distribution
```{r}
summary(aov(centile_abs.diff ~ dataset, data=mean.diffs.long))

attach(mean.diffs.long)
pairwise.t.test(x=centile_abs.diff, g=dataset, p.adj = "fdr")
detach()
```

Seems like gamlss methods are significantly better than the other combat methods and not significantly different from raw centile scores (BH corrected)

@@

No shift in abs. differences, looking at overall diff.
```{r, fig.height=7, fig.width=8}
cent.diffs2 <- combined_preds %>%
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
  arrange(dataset) %>%
  group_by(participant) %>%
  mutate(across(all_of(pheno_list), ~ . - first(.), .names = "diff_{.col}")) %>%
  ungroup()
```

@@

Test for differences in distribution
```{r}
summary(aov(z_abs.diff ~ dataset, data=mean.z.long))

attach(mean.z.long)
pairwise.t.test(x=z_abs.diff, g=dataset, p.adj = "fdr")
detach()
```

@@

Test for differences in centile offset distributions
```{r}
mean.diffs.long_filt <- mean.diffs.long %>%
  filter(!str_detect(dataset, "_refA"))

@@

attach(mean.diffs.long_filt)
pairwise.t.test(x=centile_abs.diff, g=dataset, p.adj = "fdr")
detach()
```

@@

beta.plot2
```

@@

proportion_beta_plot2
```

@@

```{r}
cent.diffs.plot2 <- mean.diffs.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  filter(!str_detect(dataset, "_refA")) %>%
  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
  ggplot() +
  geom_histogram(aes(x=centile_abs.diff, fill=dataset), color="black", alpha=0.4, position="identity") +
  theme_linedraw(base_size = 12) +
  scale_x_continuous(labels = label_comma()) + facet_wrap(~pheno_cat) +
  #paletteer::scale_fill_paletteer_d("Polychrome::light") +
  ggtitle("Mean absolute differences in centile scores across \nComBat processing pipelines for each phenotype")

cent.diffs.plot2
```

>>>>

Test for differences in distribution
```{r}
#get centile errors for each subject
cent.diffs2 <- combined_preds %>%
  dplyr::filter(!str_detect(dataset, "_refA")) %>%
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
  arrange(dataset) %>%
  group_by(participant) %>%
  mutate(across(all_of(pheno_list), ~ abs(. - first(.)), .names = "diff_{.col}")) %>%
  ungroup() %>%
  dplyr::filter(dataset != "ukb_CN") #drop ref

# cf.gmvdiff <- cent.diffs2 %>%
#   dplyr::filter(dataset == "cf") %>%
#   dplyr::select(diff_GMV)
# cf.lm.gmvdiff <- cent.diffs2 %>%
#   dplyr::filter(dataset == "cf.lm") %>%
#   dplyr::select(diff_GMV) 
# 
# t.test(cf.gmvdiff, cf.lm.gmvdiff, paired=TRUE)
```

NOTE: "missing" paired comparisons between cf and cf.lm diffs are because diff values are identical between the two methods (i.e. there is literally no difference in accuracy of centile scores between the two)

```{r, warning=FALSE, message=FALSE}
#within each phenotype, iterate paired t test between each combo of combat dfs

#initalize empty dfs to store outputs
sum.df <- data.frame(pheno = character(),
                     term = character(),
                     df = double(),
                     sumsq = double(),
                     meansq = double(),
                     statistic = double(),
                     p.value = double())
t.df <- data.frame("pheno" = character(),
                   "group1" = character(),
                   "group2" = character(),
                   "p.value" = double())
pheno_diff.list <- paste0("diff_", pheno_list)

#run tests
attach(cent.diffs2)
for (pheno in pheno_diff.list) {
  df.aov <- tidy(aov(cent.diffs2[[pheno]] ~ cent.diffs2[["dataset"]]))
  df.aov$pheno <- pheno
  sum.df <- rbind(sum.df, df.aov)

  df.pairwise.t <- tidy(pairwise.t.test(x=cent.diffs2[[pheno]], g=cent.diffs2[["dataset"]], paired = TRUE, p.adj = "none"))
  df.pairwise.t$pheno <- pheno
  t.df <- rbind(t.df, df.pairwise.t)
}
detach()
```

```{r}
#apply fdr correction
sum.df <- sum.df %>%
  dplyr::filter(term != "Residuals") %>%
  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))

summary(sum.df$p.val_fdr)

t.df <- t.df %>%
  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno)),
         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                             p.val_fdr >= 0.05 ~ FALSE),
         pheno = sub("diff_", "", pheno)) %>%
  unite(comp, c("group1", "group2")) %>%
  mutate(comp = as.factor(comp))
table(t.df$comp)

t.df %>%
  group_by(comp) %>%
   summarize(n_sig_regions = sum(sig_fdr),
            n_compare = n())
```

@@

No shift in abs. differences, looking at overall diff.

@@

Test for differences in distribution
```{r}
summary(aov(z_abs.diff ~ dataset, data=mean.z.long))
tidy(aov(z_abs.diff ~ dataset, data=mean.z.long))

attach(mean.z.long)
t.obj <- pairwise.t.test(x=z_abs.diff, g=dataset, p.adj = "fdr", paired = TRUE)
detach()
names(tidy(t.obj))

test <- tidy(t.obj) %>%
  mutate(comp = paste(group1, group2))
table(test$comp)
```

@@

Test

Test for differences in centile offset distributions
```{r}
mean.diffs.long_filt <- mean.diffs.long %>%
  filter(!str_detect(dataset, "_refA")) %>%
  filter(dataset != "ukb_CN")

@@

attach(mean.diffs.long_filt)
pairwise.t.test(x=centile_abs.diff, g=dataset, p.adj = "bonferroni")
detach()
```

@@

beta.plot2

ggsave("male_variance_beta2.jpeg", beta.plot2, width=10, height=10)
```

@@

proportion_beta_plot2
ggsave("male_variance_prop.beta2.jpeg", proportion_beta_plot2, width=10, height=10)
```

@@

```{r}
#Define scales
scales_y <- list(
  `Global Volume` = scale_y_continuous(limits = c(0, 4)),
  `Regional CT` = scale_y_continuous(limits = c(0, 70)),
  `Regional SA` = scale_y_continuous(limits = c(0, 70)),
  `Regional Volume` = scale_y_continuous(limits = c(0, 70))
)

cent.diffs.plot2 <- mean.diffs.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  filter(!str_detect(dataset, "_refA")) %>%
  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
  ggplot() +
  geom_histogram(aes(x=centile_abs.diff, fill=dataset), color="black", alpha=0.4, position="identity") +
  theme_linedraw(base_size = 12) +
  scale_x_continuous(labels = label_comma()) + 
  facet_wrap(~pheno_cat) +
  #facetscales::facet_grid_sc(pheno_cat~.,
                #scales = list(y = scales_y)) +
  #paletteer::scale_fill_paletteer_d("Polychrome::light") +
  ggtitle("Mean absolute differences in centile scores across \nComBat processing pipelines for each phenotype")

cent.diffs.plot2
#ggsave("mean_centile_diffs2.jpeg", cent.diffs.plot2, width=10, height=10)
```
